---
title: "Task 4"
author: Ignacio Almodóvar & Andrés Mejía Rodríguez
date: 18/10/21
output: pdf_document
---

# Programming in R
# Excercises 1

## Excercise 1: Create the following vectors

### a) [10,11,12,...,38]

```{r}
a=c(10:38)
```
```{r,echo=FALSE}
a
```


### b) [30,29,28,...,1]
```{r}
b=c(30:1)
```
```{r,echo=FALSE}
b
```
### c) [1,2,3,4,3,2,1]
```{r}
c=c(1:4,3:1)
```
```{r,echo=FALSE}
b
```


### d) [2,4,6,...,16,18,20]
```{r}
d=seq(2,20,2)
```
```{r,echo=FALSE}
d
```
### e) [1,2,3,...,1,2,3] 10 times repeated
```{r}
e=rep(c(1:3),10)
```
```{r,echo=FALSE}
e
```
### f) [1,2,3,...,1]
```{r}
f= e[-c(length(e),length(e)-1)]
```
```{r,echo=FALSE}
f
```
### g) ["label 1", label 2",... "label 30"]
```{r}
g=paste("label",1:30)
```
```{r,echo=FALSE}
g
```
### h) ["label-1", label-2",... "label-30"]
```{r}

h=paste('label-',1:30,sep ='')
```
```{r,echo=FALSE}
h
```
### i) $x^2e(x)$
```{r}
x=seq(0.1,1,0.1)
f=function(x){
  return(x^2*exp(x))
}
i=f(x)
```
```{r,echo=FALSE}
i
```

## Excercise 2: Calculate the following sums with and without using loops

### a) $\sum^{23} _{j=5} (j^2+3*j^{0.5})$

With loop
```{r}
aloop=0
for(i in 5:23){
  aloop=aloop+((i^2)+(3*(i^0.5)))
}
```
```{r,echo=FALSE}
aloop
```

Without loop
```{r}
f1=function(x){
  return(x^2+(3*(x^0.5)))
}

anoloop=sum(f1(5:23))
```
```{r,echo=FALSE}
anoloop
```


### b) $\sum^{18} _{i=1} (\frac{1.3^i}{i})$

With loop
```{r}
bloop=0
for(i in 1:18){
  bloop=bloop+(1.3^i)/i
}
```
```{r,echo=FALSE}
bloop
```

Without loop
```{r}
bnoloop=sum(sapply(1:18,function(x){1.3^x/x}))
```
```{r,echo=FALSE}
bnoloop
```


### c) $\sum^{10}_{i=1}\sum^{6}_{j=1} (\frac{i^4}{3+j})$

With loop
```{r}
cloop=0
for(i in 1:10){
  cloopj=0
  for(j in 1:6){
    cloopj= cloopj + (i^4)/(3+j)
  }
  cloop=cloop+ cloopj
}
```
```{r,echo=FALSE}
cloop
```

Without loop
```{r}
f3=function(x,y){
  return((x^4)/(3+y))
}

cnoloop=sum(outer(1:10,1:6,f3))
```
```{r,echo=FALSE}
cnoloop
```

## Excercise 3: 

### a) What does next code do?
```{r}
set.seed(75) 
```
This function sets the starting number used when generating a random sequence. It ensures that you always get the same sequence of random numbers each time you run the process


```{r}
M=matrix(sample(1:10,size=60,replace=TRUE),nrow=6,ncol=10) 
```
The first function matrix, builds a 6x10 matrix from the elements specified in the first parameter which in this case uses the function sample to generate a sample of a specified size (60 in this case) with numbers between 1-10 and allowing replacement.

### b) Find the number of entries in each row that are grater than 4

```{r}
counter=function(x){
  c=c(rep(0,nrow(x)))
for(i in 1:nrow(x)){
  for(j in 1:ncol(x)){
    if(x[i,j]>=4){
      c[i]=c[i]+1
    }
  }
}
  print(c)
}
counter(M)
```


### c) Replace the third column of the previous vector M by the sum of the second and third column

```{r}
M[,3]=(M[,2]+M[,3])
```
```{r,echo=FALSE}
M
```



## Excercise 4: Write a function which takes a single argument which is a matrix. The function must return a matrix which is the same as the function argument but every odd number is doubled.
```{r}
A=matrix(1:9,3,3)

doubleodd= function(x){
  for(i in 1:nrow(x)){
    for(j in 1:ncol(x)){
      if(x[i,j]%%2==0){
        x[i,j]=2*x[i,j]
      }
    }
}
return(x)
}

doubleodd(A)
```






